---
title: "DATA606_HW2"
author: "Vincent Colardi"
date: "September 19, 2018"
output: html_document
---
#### Asssignment #2 Graded: 2.6, 2.8, 2.20, 2.30, 2.38, 2.44

## Exercise 2.6 
### Dice rolls. If you roll a pair of fair dice, what is the probability of 
#### (a) getting a sum of 1? 
#### **[My Answer]** 0% Probablity
#### (b) getting a sum of 5? 
#### **[My Answer]** 4/36
#### (c) getting a sum of 12? 
#### **[My Answer]** 1/36

## Exercise 2.8
### Poverty and language. The American Community Survey is an ongoing survey that provides data every year to give communities the current information they need to plan investments and services. The 2010 American Community Survey estimates that 14.6% of Americans live below the poverty line, 20.7% speak a language other than English (foreign language) at home, and 4.2% fall into both categories.59
#### (a) Are living below the poverty line and speaking a foreign language at home disjoint? 
#### **[My Answer]** They are not disjoint. It is possible that a family could be living below the poverty line and speak a foreign language at home 
#### (b) Draw a Venn diagram summarizing the variables and their associated probabilities.
#### **[My Answer]** See Venn diagram below 
```{r}
library(VennDiagram)
draw.pairwise.venn(24.9, 18.8, 4.2, category = c("Speak Non-English at Home", "Live Below Poverty Line"), lty = rep("blank", 
    2), fill = c("light blue", "pink"), alpha = rep(0.5, 2), cat.pos = c(0, 
    0), cat.dist = rep(0.025, 2))
```

#### (c) What percent of Americans live below the poverty line and only speak English at home? 
#### **[My Answer]** 100 - 20.7 = 79.3 (Speak English), (.1460 - .0420) + .7930 = .897 or 90%
#### (d) What percent of Americans live below the poverty line or speak a foreign language at home? 
#### **[My Answer]** .2070 + 0420 = .2490 or 25%
#### (e) What percent of Americans live above the poverty line and only speak English at home? 
#### **[My Answer]** 100 - 14.6 = 85.4 P(Americans live above poverty line or speak english at home) = .8540 + .7930 - .8970 =  .75 
#### (f) Is the event that someone lives below the poverty line independent of the event that the person speaks a foreign language at home?
#### **[My Answer]** It is independent and they are disjoint events. There is a 25% chance that a person below the poverty line and speaks a foreign language at home.  

## Exercise 2.20
### Assortative mating. Assortative mating is a nonrandom mating pattern where individuals with similar genotypes and/or phenotypes mate with one another more frequently than what would be expected under a random mating pattern. Researchers studying this topic collected data on eye colors of 204 Scandinavian men and their female partners. The table below summarizes the results. For simplicity, we only include heterosexual relationships in this exercise.65

#### (a) What is the probability that a randomly chosen male respondent or his partner has blue eyes? 
#### **[My Answer]** 78 / 108 = .72
#### (b) What is the probability that a randomly chosen male respondent with blue eyes has a partner with blue eyes? 
#### **[My Answer]** 78 / 114 = .68
#### (c) What is the probability that a randomly chosen male respondent with brown eyes has a partner with blue eyes? What about the probability of a randomly chosen male respondent with green eyes having a partner with blue eyes?
#### **[My Answer]** 19 / 54 = .35
#### (d) Does it appear that the eye colors of male respondents and their partners are independent? Explain your reasoning.
#### **[My Answer]** There does not appear to be any compelling evidence in the outcomes, that men & women with the same eye color choose each other as mates; the results are 78/114 or 68% (Blue), 23/54 and 43% (Brown) and 16/36 or 44% (Green)

## Exercise 2.30
### Books on a bookshelf. The table below shows the distribution of books on a bookcase based on whether they are non???ction or ???ction and hardcover or paperback.

#### (a) Find the probability of drawing a hardcover book First then a paperback fiction book second when drawing without replacement. 
#### **[My Answer]** 28/95 * 59/94 = .1850

#### (b) Determine the probability of drawing a fiction book first and then a hardcover book second, when drawing without replacement. 
#### **[My Answer]** 72/95 * 28/94 = .2258 

#### (c) Calculate the probability of the scenario in part (b), except this time complete the calculations under the scenario where the first book is placed back on the bookcase before randomly drawing the second book.
#### **[My Answer]** 72/95 * 28/95 = .2234

#### (d) The final answers to parts (b) and (c) are very similar. Explain why this is the case.
#### **[My Answer]** Replacing the book only adds 1 to the denominator of the second probablity which does not drastically impact the overall result of the calculation

## Exercise 2.38
### Baggage fees. An airline charges the following baggage fees: $25 for the First bag and $35 for the second. Suppose 54% of passengers have no checked luggage, 34% have one piece of checked luggage and 12% have two pieces. We suppose a negligible portion of people check more than two bags.

#### (a) Build a probability model, compute the average revenue per passenger, and compute the corresponding standard deviation. 

#### **[My Answer]** E(X) = 0 * P(X = 0) + 35 * P(x = 35) + 60 * P(X = 60)
#### E(X) = 0 * .54 + 35 * .34 + 60 * .12, **E(X) = 19.10 

#### Table Row Legend
#### Formula 1 $x_{i}$                          
#### Formula 2 P(X=$x_{i}$)                     
#### Formula 3 $x_{i}$ x P($x_{i}$)                  
#### Formula 4 $x_{i}$ - $\mu$                  
#### Formula 5 ($x_{i}$-$\mu)^{2}$                  
#### Formula 6 (x-$x_{i})^{2}$ x P(x=$x_{i}$)    

#### Standard Deviation = Square root of 483.74 = 22


```{r}

b <- matrix (c(0, 0.54, 0, -19.10,364.81, 197,
          35, 0.34, 11.9, 15.9, 252.81, 86,
          60, 0.12, 7.2, 40.9, 1672.81, 200.74,
          0.0, 0.0, 0.0, 0.0, 0.0, 483.74), nrow=6, ncol=4)

baggage <- matrix (c(0, 35, 60, 0,
                    0.54, 0.34, 0.12, 0.0,
                    0, 11,9, 7.2, 0.0,
                    -19.10, 15.9, 40.9, 0.0,
                    364.81, 252.81, 1672.81, 0.0,
                    197, 86, 200.74, 483.74), nrow=6, ncol=4)
colnames(b) <- c('1','2','3','Total')
rownames(b) <- c('Formula 1', 'Formula 2','Formula 3','Formula 4','Formula 5','Formula 6')
b.table <- as.table(b)
b.table
```


#### (b) About how much revenue should the airline expect for a flight of 120 passengers? With what standard deviation? Note any assumptions you make and if you think they are justified
#### **[My Answer]** The standard deviation should be $22 which is based on the expected value calculated in (a)
```{r}
flight_rev = ((.54 * 120) * 0) + ((.34 * 120) * 25) + ((.12 * 120) * 60) 
flight_rev
```

## Exercise 2.44

### Income and gender. The relative frequency table below displays the distribution of annual total personal income (in 2009 in???ation-adjusted dollars) for a representative sample of 96,420,486 Americans. These data come from the American Community Survey for 2005-2009. This sample is comprised of 59% males and 41% females.

#### (a) Describe the distribution of total personal income. 
#### **[My Answer]**  The distribution is symmetric, unimodal distribution

#### (b) What is the probability that a randomly chosen US resident makes less than $50,000 per year? 
#### **[My Answer]** 21.2 + 18.3 + 15.8 + 4.7 + 2.2 = 62.2%

#### (c) What is the probability that a randomly chosen US resident makes less than $50,000 per year and is female? Note any assumptions you make. 
#### **[My Answer]** (.41 * 96420486) * .62) / (.41 * 96,420,486) = 39%; assumption is that the income figures are not gender specific. 

#### (d) The same data source indicates that 71.8% of females make less than $50,000 per year. Use this value to determine whether or not the assumption you made in part (c) is valid.
#### **[My Answer]** The assumption is way off; 39% compared to 71.8%
